### 2021-07-23 scripts/DependencePlots.Rmd

---
title: "DependencePlots"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(semtree)
library(metaforest)
library(gridExtra)
library(grid)
```

```{r loading objects}
forestdata <- readRDS("data/forestfile.RData")
vim <- readRDS("data/forestvim.RData")
```

```{r extracting importance values}
sort.values = T 
aggregate = "mean" 
horiz = T
las = 1 
convergence = F 
scale = "absolute" 
xlim = NULL 
head = NULL 
tail = NULL 
na.omit = FALSE

vimp <- vim
x <- semtree:::aggregateVarimp(vimp, aggregate, scale, na.omit)

vnames <- vimp$var.names
low <- min(x, na.rm = T) - 1
filt <- is.na(x)
x[filt] <- low
srt <- sort(x, index.return = T)
x <- x[srt$ix]
vnames <- vnames[srt$ix]
x[x <= (low + 0.5)] <- NA

selection <- 1:length(x)

impmat <- x[selection]
```

```{r partialDependence residual}
indxvec <- matrix(c(1:5, (length(impmat)-4):length(impmat)), nrow=1)

pDres <- readRDS("data/pDres.RData")
```

```{r partialDependence meani}
pDmi <- readRDS("data/pDmi.RData")
```

```{r partialDependence means}
pDms <- readRDS("data/pDms.RData")
```

```{r partialDependence meanq}
pDmq <- readRDS("data/pDmq.RData")
```

```{r plots,  results="asis"}
texttempl <- "This is variable of importance rank %i
Variable: %s
"
plotres <- list()
plotmi <- list()
plotms <- list()
plotmq <- list()

for(i in 1:length(indxvec)){
  plotres[[i]] <- ggplot() +
  geom_line(mapping=aes(x=as.matrix(pDres[[i]]$xgrid), y=as.matrix(unlist(pDres[[i]]$dict))), col="red") +
  geom_point(mapping=aes(x=as.matrix(pDres[[i]]$xgrid), y=as.matrix(unlist(pDres[[i]]$dict))), col="black", 
             shape=3, size=3) +
  labs(x=names(impmat)[indxvec[i]], y="residual") +
  theme_minimal()

  plotmi[[i]] <- ggplot() +
  geom_line(mapping=aes(x=as.matrix(pDmi[[i]]$xgrid), y=as.matrix(unlist(pDmi[[i]]$dict))), col="red") +
  geom_point(mapping=aes(x=as.matrix(pDmi[[i]]$xgrid), y=as.matrix(unlist(pDmi[[i]]$dict))), col="black", 
             shape=3, size=3) +
  labs(x=names(impmat)[indxvec[i]], y="mean i") +
  theme_minimal()

  plotms[[i]] <- ggplot() +
  geom_line(mapping=aes(x=as.matrix(pDms[[i]]$xgrid), y=as.matrix(unlist(pDms[[i]]$dict))), col="red") +
  geom_point(mapping=aes(x=as.matrix(pDms[[i]]$xgrid), y=as.matrix(unlist(pDms[[i]]$dict))), col="black", 
             shape=3, size=3) +
  labs(x=names(impmat)[indxvec[i]], y="mean s") +
  theme_minimal()

  plotmq[[i]] <- ggplot() +
  geom_line(mapping=aes(x=as.matrix(pDmq[[i]]$xgrid), y=as.matrix(unlist(pDmq[[i]]$dict))), col="red") +
  geom_point(mapping=aes(x=as.matrix(pDmq[[i]]$xgrid), y=as.matrix(unlist(pDmq[[i]]$dict))), col="black", 
             shape=3, size=3) +
  labs(x=names(impmat)[indxvec[i]], y="mean q") +
  theme_minimal()
}

for(i in 1:length(indxvec)){
  grid.arrange(plotres[[i]], plotmi[[i]], plotms[[i]], plotmq[[i]], nrow=2,
               top=textGrob(sprintf(texttempl, (length(impmat)-indxvec[i]+1), names(impmat)[indxvec[i]])))
}
```