### 2023-11-24 presentation_copenhagen2023.Rmd
update 2023-11-24 10:41:01.673245

---
title: "Machine Learning can Advance Theory Formation in the Social Sciences"
author: "Caspar J. Van Lissa<sup>1</sup>"
date: "<sup>1</sup>: Tilburg University, dept. Methodology & Statistics<br>Funded by NWO Veni grant  VI.Veni.191G.090"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
output: 
  ioslides_presentation:
    widescreen: true
    css: slide.css
bibliography      : ["manuscript/references.bib", "presentation/all_refs.bib"]
---

```{r setup, include=FALSE}
#css: https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Conceptual introduction

## Central thesis

_Machine learning_ can help advance _theory formation_ in the social sciences

* Replication crisis [@scheelWhyMostPsychological2022; @lavelleWhenCrisisBecomes2021]
* Current solutions improve deductive research (theory-testing)
    + questionable research practices [@wichertsDegreesFreedomPlanning2016]
    + replication
    + preregistration

Many preregistered hypotheses are not supported (56%) [@scheelExcessPositiveResults2021]

* Lack of "good theories" another explanation for replication crisis

## Defining theory

> **Theory**: A (causal?) explanation of phenomena reliably evidenced by data [@bogenSavingPhenomena1988]

* Allows for derivation of testable hypotheses for deductive research

## Characteristics of Good Theory

@meehlAppraisingAmendingTheories1990

1. phenomena to be explained
2. putative causal connections between them
3. direction of the effect
4. relative importance of predictors
5. interaction effects
6. functional form of associations

Most social scientific theories are [@scheelWhyMostPsychological2022]:

* insufficiently precise
* too flexible
* But cf. @walasekEvolutionaryModelSensitive2022

## Why do we need theory?

* Social science strives to understand, not only predict
* Useful as an instrument of cumulative knowledge acquisition
* Communicating social scientific research to its consumers

## Deduction vs induction

```{r, out.width="60%"}
library(tidySEM)
lay <- get_layout("the1", "the2",
                  "hyp1", "hyp2",
                  "pat1", "obs2",
                  "obs1", "con2", rows = 4)
nodes <- data.frame(
  name = as.vector(lay),
  label = c("Theory", "Working hypothesis", "Pattern", "Observation", "Theory", "Hypothesis", "Observation", "Test"),
  shape = rep("rect", length(lay)),
  group = rep(c("Inductive", "Deductive"), each = 4)
  )
edges <- data.frame(
  from = c("the1", "hyp1", "pat1", "the2", "hyp2", "obs2"),
  to = c("hyp1", "pat1", "obs1", "hyp2", "obs2", "con2"),
  arrow = rep(c("first", "last"), each = 3),
  group = rep(c("Inductive", "Deductive"), each = 3)
)
g <- prepare_graph(edges = edges, layout = lay, nodes = nodes)
plot(g)
```

* Lack of theory cannot be overcome by improving deductive research
* Theory formation requires inductive (exploratory) research [@creswellDesigningConductingMixed2017]

## Empirical cycle

```{r, out.width="40%"}
knitr::include_graphics("presentation/empiricalcycle.png")
```

De Groot’s empirical cycle

* Model of cumulative knowledge acquisition through scientific research [@degrootMethodologieGrondslagenVan1961]

## Rigorous exploration

How can we improve exploratory research?

* Confirmatory methods poorly suited to exploratory research
    + p-values: Type I error is 5%, but false discovery rate (FDR) is 14-50% [@vidgenPValuesMisunderstoodMisused2016].
    + Model fit indices: manually specify models, no guarantee that best model is included
* Role of flexibility: there are infinite ways to explore
* "unguided exploration" is effortful and inflates the risk of spurious results

## Machine learning

**Machine learning**:

* Automated model building
* Checks and balances to prevent overfitting
* Maximize predictive performance and generalizability of results

Psychology can benefit from its superior predictive performance [@yarkoniChoosingPredictionExplanation2017]

**NEW**: Implications for the theory crisis have not yet been discussed

## Machine learning vs unguided exploration

```{r, out.width="100%"}
knitr::include_graphics("Figure Machine learning 4.png")
```

# Advantages of Machine Learning

## Phenomena detection

First step in *Theory Construction Methods* is identifying relevant phenomena [@borsboomTheoryConstructionMethodology2020]
*
* Few quantitative methods for phenomena detection (aside from expert opinion)
* Phenomena: stable and general features of the world
* **Cross-validated patterns of association** may constitute phenomena
* **Text mining** existing literature may reveal what phenomena other authors have identified [@vanlissaMappingPhenomenaRelevant2021]

## Holistic approach

Every study examines only a piece of the puzzle; we never see the complete picture

Machine learning accommodates more predictors than classical methods

* Regularization 
* Variable selection

Including all relevant predictors is important:

* Assumption for causal interpretation
    + But mind multicolinearity and collider bias
* Good theory incorporates most important causes
* Include potential predictors from various theories, and undertheorized factors
    + Assess the relative utility across studies
    + Cannot do this across studies

## Complex effects

Many machine learning methods accommodate:

* non-linear effects 
* higher-order interactions, without having to specify the nature of these effects a-priori.
* Automatic: tree-based methods
* Manual: penalized methods

Social scientific theories rarely account for complex effects

* Machine learning can provide nuance by revealing them

## Theoretical elements

Some methods allow for theory guided exploration using machine learning

When (part of) the theory can be expressed as a structural equation model, and part still remains to be discovered through exploratory research

* E.g., some relevant constructs are latent variables
* E.g., development follows a latent growth curve (LGC)
* SEM forests [@brandmaierTheoryguidedExplorationStructural2016]
* regularized SEM [@jacobucciRegularizedStructuralEquation2016], (Ernst & Peikert 2022)

When the theory is a (nomological) network:

* LASSO-penalized Gaussian graphical model (GGM) [@epskampGeneralizedNetworkPsychometrics2017]
* E.g.: network theory of major depression [@cramerMajorDepressionComplex2016]

## Person-centered approaches

Call for person-centered research [@laursenPersonCenteredVariableCenteredApproaches2006]

Explain heterogeneity at a more fine-grained level than the whole sample

* Latent class analyses
    + "Which individuals are similar?"
    + Other *unsupervised learning* methods exist
* Multilevel models
    + Allow for, but rarely explain heterogeneity
    + Includes RI-CLPM and DSEM
    + Could use regularized SEM or SEM forests
* Tree-based models
    + Group individuals based on predictors to maximize homogeneity of the outcome
    + "Why are these individuals similar?"

## Generalizability

When using data models to guide theory formation, generalizability is essential [@haslbeckModelingPsychopathologyData2021]

ML: Checks and balances to curtail *overfitting*

* cross-validation to estimate predictive accuracy

<!-- Any model captures both *signal* and *noise* -->

<!-- **Overfitting**: model captures so much noise that it generalizes poorly to new samples -->

```{r, out.width="40%"}
library(ggplot2)
colors <- c("Variance" = "red", "Bias" = "green", "Total error" = "black")

ggplot() +
  xlim(-3, 3) +
geom_function(
  data = NULL,
  stat = "function",
  position = "identity",
  fun = ~.x^2,
  na.rm = FALSE,
  inherit.aes = TRUE,
  linewidth = 1,
  aes(colour = "Total error")
) +
  geom_function(
    data = NULL,
    stat = "function",
    position = "identity",
    fun = ~2^.x -2,
    na.rm = FALSE,
    inherit.aes = TRUE,
    aes(colour = "Bias"),
    linewidth = 1
  )+
  geom_function(
    data = NULL,
    stat = "function",
    position = "identity",
    fun = ~2^-.x -2,
    na.rm = FALSE,
    inherit.aes = TRUE,
    aes(colour = "Variance"),
    linewidth = 1
  ) +
  labs(x = "Model complexity",
       y = "Optimal model complexity",
       color = "Legend") +
  scale_color_manual(values = colors) +
  geom_vline(xintercept = 0, linetype = 2) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank())
```

## From models to theory

Naive interpretative approach

* Start with "interpretable machine learning" metrics [@henningerInterpretableMachineLearning2022]

* Variable importance
    + congruent with theoretical assumptions about important predictors?
    + any theoretically important predictors rank low?
    + any undertheorized factors rank high?
* Marginal associations
    + non-linear effects?
    + high importance but flat marginal association?
* Exemplar-based interpretation

## From models to theory 2

Data models -> formal theory [@haslbeckModelingPsychopathologyData2021]

* Abductive Formal Theory Construction (AFTC) framework
    + construct formal theory
    + Generate data
    + Mathematically compare to empirical data
    + like a posterior predictive check in Bayes
    + Discrepancies: amend formal theory
* "Light" version of AFTC: Comparing predictive performance of simpler parametric model based on "interpretation" to more flexible machine learning model

<!-- ## Summary -->

<!-- * There is a paucity of good theory -->
<!-- * Need for exploratory research for theory formation -->
<!-- * Machine learning for rigorous exploration -->
<!--     + automates model building -->
<!--     + incorporates checks and balances for generalizable results -->
<!-- * Unsupervised learning can assist in phenomenon detection -->
<!-- * Supervised learning to identify important predictors -->
<!-- * Some algorithms incorporate basic theoretical elements -->

# Example: Text Mining Systematic Review

## Emotion regulation in adolescence

Developmentally sensitive period [@zimmermannEmotionRegulationEarly2014]

20% develop psychopathology [@leeAdolescentMentalHealth2014]

Potentially lifelong implications for mental health and well-being

```{r, out.height="20%"}
knitr::include_graphics("trajectories.jpg")
```

## Knowledge gap

Substantial empirical research, but:

* Different (sub)disciplines have approached the topic in disparate ways [@riedigerEmotionRegulationAdolescence2014]
* No consistent terminology [@bariolaChildAdolescentEmotion2011]
* No conceptual frameworks [@stifterEmotionRegulation2019]
* No overarching theoretical framework [@bussTheoriesEmotionalDevelopment2019]

Knowledge must be consolidated into overarching theory

## Limitations of relevant theory

* Few explicitly address adolescence
* Few comprehensively address predictors of emotion regulation
* None directly guide contemporary research 
* Theories vary widely in scope: Some are broad and non-specific; others describe a specific phenomenon in detail, but lack a broader perspective.
    + Broad theories can frame any research, specific theories generate hypotheses.
    + It would be beneficial to bridge these levels of analysis.

## Towards integrative theory

* First step: identifying relevant phenomena [@borsboomTheoryConstructionMethodology2020]
    + Often done by systematic review
* Narrative reviews: small samples, confirmation bias, emphasize positive results [@littellEvidencebasedBiasedQuality2008]
* Introducing text mining systematic review (TMSR) [@vanlissaMappingPhenomenaRelevant2021]:
    + Unlimited sample size, transparent, objective, reproducible

6653 papers on **emotion regulation** in **adolescence** [age 10-24]

<https://doi.org/10.1007/s40894-021-00160-7>

## Open science

All code and data available at <https://github.com/cjvanlissa/veni_sysrev>

Workflow for Open Reproducible Code in Science (WORCS) used to make analyses reproducible [@vanlissaWORCSWorkflowOpen2020]

## Baseline

Manual selection of key constructs from relevant theory and prior reviews [ @bariolaChildAdolescentEmotion2011; @coe-odessEmergentEmotionsAdolescence2019]:

* Bioecological model [@bronfenbrennerBioecologicalModelHuman2007]
* Transactional model [@sameroffUnifiedTheoryDevelopment2010]
* Hall's notion of "storm and stress"
* Theory of normative emotional development [@sroufeEmotionalDevelopmentOrganization1995]
* Tripartite model [@morrisRoleFamilyContext2007]
* Internalization model [@holodynskiDevelopmentEmotionsEmotion2006]
* Polyvagal theory [@porgesOrientingDefensiveWorld1995].
* Model of social-affective engagement and goal flexibility [@croneUnderstandingAdolescencePeriod2012]
* Process model of emotion regulation [@grossHandbookEmotionRegulation2013]
* Social Information Processing Theory [@lemeriseIntegratedModelEmotion2000]

## Baseline network {.smaller}

```{r figbaseline, warning=FALSE, eval = TRUE, results='asis', out.width = "100%", fig.cap="Phenomena relevant to adolescents' emotion regulation according to theory (a) and narrative reviews (b; transparent nodes indicate constructs also present in the theory)."}
knitr::include_graphics("presentation/baseline_network.png") 
```

(a) Theory (b) narrative reviews; transparent nodes indicate constructs also present in theory.

## {.fullslide}

![](presentation/varimps_small.png)

## Co-occurrence graph {.smaller}

```{r networks1, eval = TRUE, results='asis', out.width = "100%"}
knitr::include_graphics("presentation/tmnetworks_small.png")
```

Author keywords (a) and abstracts (b)

## Results

* Both analyses reflected some constructs from theoretical literature
    + Particularly related to neurodevelopment and socialization
* Mental health-related outcomes feature prominently
    + Emotion regulation implicated in mental health problems [@leeAdolescentMentalHealth2014],
    + This underlines the importance of research in this area
* Substantial correspondence between keywords and abstracts, suggesting validity of method
* Networks are sparse; few connections among constructs: fragmented literature

## Undertheorized themes

* Developmental disorders
* Physical health (sic)
* External stressors
* Structural disadvantage
* Addictive behavior
* Identity and moral development
* Sexual development

## Implications

* Empirical research relies on theory; under-theorized phenomena may be overlooked
    + TMSR offers guidance for phenomena to consider as confounders/causes
* Unembedded terms indicate promising areas of future research
* TMSR network can serve as proto-theoretical nomological network

## Next Steps

Rasoul Norouzi is further developing this method

* Goal: Reconstructing a causal loop diagram based on authors' causal claims in literature

Together with Denny Borsboom, applying TMSR v2 to Cooperation Databank

* Using sciBERT vector embeddings and HDBSCAN clustering instead of dictionary

# Example: Predictors of emotion regulation development

## Person-centered models

```{r, out.height="50%"}
knitr::include_graphics("groups.png")
```

* Predictors of latent growth model?
  + Only linear differences, few predictors (power)
* Multigroup latent growth model?
  + Non-linear differences, but only one moderator
* Latent class growth model with auxilliary variables?
  + Groups with regard to trajectory, not predictors

## Method

SEM-forest [@brandmaierTheoryguidedExplorationStructural2016]

1. Bootstrap sample
2. On each sample, estimate a SEM-model
3. Consider $k$ candidate predictors
4. Identify predictor and value that maximizes $LR$ of post-split multi-group model
5. Average predictions across bootstrap samples

We used 1000 bootstrap samples, $k = 9$

## SEM Forest

```{r, out.width="70%"}
knitr::include_graphics("tree.png")
```

<!-- TS: Confirmatory versus exploratory research: Define the distinction. -->
<!-- TS: Some longitudinal research falls between these categories. For example, complex longitudinal studies, where prior research is sufficient to lead us to hypothesize *some* associations between two constructs, but not sufficiently specific to let us hypothesize which parameters will be significant. -->

<!-- ## Results -->

<!-- * Both analyses reflected some constructs from theoretical literature -->
<!--     + neurodevelopment, socialization -->
<!-- * Undertheorized themes -->
<!--     + Developmental disorders -->
<!--     + Physical health (sic) -->
<!--     + External stressors -->
<!--     + Structural disadvantage -->
<!--     + Addictive behavior -->
<!--     + Identity and moral development -->
<!--     + Sexual development -->

<!-- * This offers guidance for phenomena to consider as relevant predictors -->

## Study design

RQ1: What are the most important predictors of adolescents' trajectories of emotion regulation development?

RQ2: What is the nature of the association of the predictors with the trajectories?

**Participants**

* RADAR data (https://www.uu.nl/en/research/radar)
* 497 Dutch adolescents (283 boys; age at T1: M =13.03, SD = 0.46)
* Collected between 2006-2011
* Most families were  medium- to high-SES (10% low-SES)

## Candidate predictors

Predictors: 87

Categories:

* Demographics (e.g., SES)
* Biological (age, sex, pubertal development)
* Individual differences (big5, BIS/BAS, empathy)
* Risk behavior (in/externalizing symptoms, substance use, delinquency, school performance, victimization)
* Relationship quality (parents/peers)
* Parenting (perceived and parent-reported)
* Conflict resolution styles

## Dependent variable

Difficulties in emotion regulation [@gratzMultidimensionalAssessmentEmotion2004]

* 24 items (omitted 'emotional awareness' subscale)
* E.g.: “When I’m upset, I have difficulty thinking about anything else”
* 1 (“Almost never”) to 5 (“Almost always”)
* Excellent reliability, $(.94 > \alpha > .95)$.
* Metric invariance across waves
* Quadratic growth model had good fit
  + CFI = .996, TLI = .993, RMSEA = .047, SRMR = 0.020

## Open science

Workflow for Open Reproducible Code in Science (WORCS) used to make analyses reproducible [@vanlissaWORCSWorkflowOpen2020]

All code and synthetic data available at

<https://github.com/cjvanlissa/veni_forest>

# Results

## Variable importance

```{r, out.width="90%"}
knitr::include_graphics("pcomb2.png")
```

## Variable importance 2

```{r, out.width="100%"}
knitr::include_graphics("violin.png")
```

## Marginal association

```{r, out.width="40%"}
knitr::include_graphics("plots/pdp1-12.png")
```

## Summary

Important:

* Best predictors: Personality and related constructs
* Best parental predictor is autonomy support (bal. rel.)
  + Negative/intrusive parenting more predictive than other positive practices
* Conflict frequency and behavior
* Empathy, particularly personal distress

Less important than expected:

* SES
* Bullying/victimization
* Delinquency
* Substance use
* Monitoring

## Reflection content

* Proximal factors more predictive of ER development than distal predictors
    + Order mirrors bioecological model
* Most important predictors are routinely assessed (big 5) or overt (conflict)
  + Prime candidates for early risk assessment
  + Conflict resolution behavior can be taught
    - Target for intervention if association is causal
* Emphasis on parenting (in literature and theory) might not be justified
  + Congruent with RI-CLPM showing few parenting effects in adolescence [@vanlissaRoleFathersMothers2019]

## Reflections form

* Many predictors show non-linear effects: Emotion dysregulation only for +/-1SD
* Some predictors show almost no marginal effect (e.g., father's age, drug use)
  + This suggests they might be important in **interactions**

# Example: Predictors of Infection Prevention Behavior

## Context

Prior to vaccines, infection prevention behaviors were the only defense against COVID

Behavioral science could thus inform policy

Existing theories suggest a limited set of relevant predictors

We used ML to cast a wider net and identify important predictors at the individual and country level

## Study design

**Participants**

* PsyCorona consortium
* 56.072 participants across 28 countries
* Collected between March-May 2020

## Variables

* DV: Self-reported infection prevention behaviors (handwashing, avoiding crowds, isolation)
* IVs: 80 individual-level variables (16 multi-item composites)
* Enriched with 35 country-level variables
    + E.g., political stability, government effectiveness
    + Some matched to date of participation (pandemic severity, government response)

## Open science

Workflow for Open Reproducible Code in Science (WORCS) used to make analyses reproducible [@vanlissaWORCSWorkflowOpen2020]

All code and synthetic data available at

<https://github.com/cjvanlissa/covid19_metadata>

# Results

## Variable importance

```{r, out.height="100%"}
knitr::include_graphics("presentation/varimp_covid.png")
```

## Summary

* Injunctive norms most important
    + Consistent with theory of planned behavior
* Coping styles were important:
    + Problem-focused coping, focus on the present/ future, hopefulness
* Five of most important predictors were country-level variables
    + E.g., national healthcare infrastructure

## Notable absences

* Perceived likelihood of infection
    + Contradicts Health Belief Model
* Psychological and material struggle
    + E.g., loneliness, boredom, job insecurity
* No evidence for "intergenerational conflict of interest"
    + Confirmed by preregistered studies

## Strengths

This sample included a representative sub-sample

* Allowed us to estimate generalizability
* Predictive $R^2$ in the representative subsample was even higher than in the total hold-out sample ($R^2_{rep} = .59$, $R^2_{test} = .52$)

# Limitations and future directions

## Limitations

1. Theories are still verbal and flexible
    * Even if ML leads to new insights, it is not clear how theories should be updated
2. Ignoring causality limits interpretability of machine learning
    * Incorrect direction of causality: meaningless results
    * Conditioning on a collider: biased results

## Future direction 1: Formal theoretical object

Formal theories are expressed in mathematical terms

* unambiguous
* logically sound
* explicates assumptions
* reveals ambiguities and gaps in knowledge
* facilitates conveying ideas to others
* applying rules of formal logic implies testable hypotheses
* refutation of hypotheses dictates specific modifications

## Recall: Characteristics of Good Theory

@meehlAppraisingAmendingTheories1990

1. phenomena to be explained
2. putative causal connections between them
3. direction of the effect
4. relative importance of predictors
5. interaction effects
6. functional form of associations

## Formal Theory

* 1-2 can be described by a causal network (DAG), 3-6 refer to the nature of associations within that network
* Augmented causal networks (ACN, see Crielaard et al., 2022)

* encode both causal relationships and the nature of those relationships
* imply testable hypotheses
* shared in plain text, e.g., YAML or RDF
* version controlled
* shared conform FAIR software principles,
to be used and updated through distributed collaboration

Working on this with Aaron Peikert, Andreas Brandmaier and Felix Schönbrodt

## Learning a causal network

For example, using TMSR, based on what previous authors have written

Working on this with my PhD student, Rasoul Norouzi

## Epistemology

Working on an epistemology of machine learning-informed theory formation

With Denny Borsboom and James Woodward

## References {.smaller}

<font size = 4>
**Van Lissa, C. J. (2022). Developmental data science: How machine learning can advance theory formation in Developmental Psychology. Infant and Child Development. https://doi.org/10.1002/icd.2370**

Van Lissa, C. J. (2022). Complementing preregistered confirmatory analyses with rigorous, reproducible exploration using machine learning. Religion, Brain & Behavior. https://doi.org/10.1080/2153599X.2022.2070254

Van Lissa, C. J., Beinhauer, L., Branje, S., & Meeus, W. (2022). Using machine learning to identify early predictors of adolescent emotion regulation development. Journal of Research on Adolescence. https://doi.org/10.31234/osf.io/rbqf9

Van Lissa, C. J. (2021). Mapping Phenomena Relevant to Adolescent Emotion Regulation: A Text-Mining Systematic Review. Adolescent Research Review. https://doi.org/10.1007/s40894-021-00160-7

Van Lissa, C. J., et al., (2022). Using machine learning to identify important predictors of COVID-19 infection prevention behaviors during the early phase of the pandemic. Patterns, 3(4), 100482. https://doi.org/10.1016/j.patter.2022.100482

Douglas, K., Sutton, R. M., Van Lissa, C. J., ... & N Pontus, L. (2023). Identifying Important Individual-and Country-Level Predictors of Conspiracy Theorizing: A Machine Learning Analysis. European Journal of Social Psychology. DOI: 10.1002/ejsp.2968

</font>

## More refs